Comparison of Terrain Following and Cut Cell Grids using a Non-Hydrostatic Model

### 2015-09-17 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
clarify cell centres are centroids

@@

The model is operated in a two-dimensional $x$--$z$ slice configuration without Coriolis forces.  The model uses the C-grid staggering in the horizontal and the Lorenz staggering in the vertical such that $\theta$, $\rho$ and $\Pi$ are stored at cell centres and the covariant component of velocity at cell faces.
A minor change has been made to the non-orthogonal formulation as described in \citet{weller-shahrokhi2014} to improve stability on non-orthogonal grid, and this is detailed in appendix A.

@@

\subsection{Terrain following advection}
In the horizontal advection test, results were more accurate where the flow was aligned with the grid layers, and distortions in the BTF grid led to increased errors.  To determine the primary source of numerical error, we formulate a new tracer advection test in which the velocity field is everywhere tangential to the basic terrain following coordinate surfaces.  Misalignment on the SLEVE and cut cell grids is designed to challenge the advection scheme.

@@

\citet{good2014} found the maximum vertical velocity in their cut cell model was \SI{1e-12}{\meter\per\second}, which is better than any result obtained using the model by \citet{weller-shahrokhi2014}.  It is worth noting that, in the model used by \citet{good2014}, cell centres are in the centre of the uncut cell, resulting in the centre of some cut cells being below the ground (S.-J. Lock 2014, personal communication).  This means that the grid is effectively regular when calculating horizontal and vertical gradients.  This would account for the very small velocities found by \citet{good2014}.

>>>>

% Every term in the discretised equations of motion is a potential source of numerical error.
% Errors in calculating the horizontal pressure gradient generate spurious circulations and so much effort has been put into improving this calculation.  This is often achieved by making the grid more orthogonal, flattening terrain following coordinate surfaces [citations] or using a cut cell approach [citations].
% TF citations: mahrer1984, schaer2002, klemp2011
% cut cell citations: good2014, yamazaki2015 puts this clearly but not yet published
% In a test of a stably stratified atmosphere initially at rest, we find that spurious velocities are no more than \SI{0.1}{\meter\per\second} on any grid. [TODO: and this is much better than klemp2011]
% Advection terms also contribute to numerical error.  We show that advection is most accurate when the flow is aligned with the grid layers.  Near the surface, flow tends to follow the terrain which results in larger errors when more orthogonal grids are used.
% In a test of orographically induced gravity waves, we find that the advection of potential temperature generates grid-scale errors on the cut cell grid.  The errors persist because the Lorenz staggering of variables renders them invisible to the model. [TODO: I don't think we have enough evidence to support this final sentence]

@@

The model is operated in a two-dimensional $x$--$z$ slice configuration without Coriolis forces.  The model uses the C-grid staggering in the horizontal and the Lorenz staggering in the vertical such that $\theta$, $\rho$ and $\Pi$ are stored at cell centroids and the covariant component of velocity at cell faces.
A minor change has been made to the non-orthogonal formulation as described in \citet{weller-shahrokhi2014} to improve stability on non-orthogonal grid, and this is detailed in appendix A.

@@

\subsection{Terrain following advection}
\TODO{make it clearer that we're trying to determine if non-orthogonality or misalignment of the flow is the primary source of error}

In the horizontal advection test, results were more accurate where the flow was aligned with the grid layers, and distortions in the BTF grid led to increased errors.  To determine the primary source of numerical error, we formulate a new tracer advection test in which the velocity field is everywhere tangential to the basic terrain following coordinate surfaces.  Misalignment on the SLEVE and cut cell grids is designed to challenge the advection scheme.

@@

\citet{good2014} found the maximum vertical velocity in their cut cell model was \SI{1e-12}{\meter\per\second}, which is better than any result obtained here.  It is worth noting that the model by \citet{weller-shahrokhi2014} stores values at the geometric centre of cut cells.  However, in the model used by \citet{good2014}, cell centres are in the centre of the uncut cell, resulting in the centre of some cut cells being below the ground (S.-J. Lock 2014, personal communication).  This means that the grid is effectively regular when calculating horizontal and vertical gradients.  This would account for the very small velocities found by \citet{good2014}.
